---
title: "Lab 1"
author: "Dastin Scott"
date: "`r format(Sys.Date())`"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Task 1

```{r}
getwd()
```

# Task 2

```{r}
ddt <- read.csv("DDT.csv")
head(ddt)
```

# Task 3

**What are the qualitative variables in "ddt"? ** 

"RIVER and "SPECIES" are the qualitative variable because they aren't numeric.


**What are the quantitative variables in "ddt"?**

"MILE", "LENGTH", "WEIGHT", and "DDT" are the quantitative variables because they are numeric. 


**How many SPECIES are in the ddt data set?**
```{r}
length(unique(ddt$SPECIES))
```

**Subset the ddt data set so that we see only the observations with SPECIES, LMBASS and WEIGHT bigger than 800gms **
```{r}
subset_LMBASS_800 <- subset(ddt, SPECIES == "LMBASS" & WEIGHT > 800)
subset_LMBASS_800
```

**Subset the ddt data set so that we see only those observations where the RIVER is SCM and DDT bigger than 4.0**
```{r}
subset_SCM_DDT4 <- subset(ddt, RIVER == "SCM" & DDT > 4.0)
subset_SCM_DDT4
```

# Task 4

**Q.1 Find the mean length of fish in the ddt data frame!**

```{r}
  mean(ddt$LENGTH)
```

  Ans. `r mean(ddt$LENGTH)`

**Q.2 What is the standard deviation of the weight of fish in the ddt data frame?**

```{r}
  sd(ddt$WEIGHT)
```

  Ans. `r sd(ddt$WEIGHT)`

**Q3. Is this the plot of LENGTH vs WEIGHT from the ddt data frame?**

  Ans. No, the plot given was WEIGHT vs LENGTH, the plot below is LENGTH vs WEIGHT
  
```{r}
plot(LENGTH ~ WEIGHT, data = ddt, main = "LENGTH vs WEIGHT")
```


Q4. If v=1:20 what is the last value of v/20? 

```{r}
v <- 1:20
tail(v/20, 1)
```

  Ans. `r tail(v/20, 1)`

# Task 5

```{r}
# table of rivers
tab_river <- with(ddt, table(RIVER))
tab_river

# barplot of rivers
barplot(tab_river, main = "Counts by River", xlab = "RIVER", ylab = "Count", col = 1:length(tab_river))

# rivers x species
tab_river_species <- with(ddt, table(RIVER, SPECIES))
tab_river_species


# barplot rivers x species
library(ggplot2)
ggplot(ddt, aes(x = RIVER, fill = SPECIES)) +
  geom_bar(position = "dodge") +
  labs(title = "Rivers crossed with Species",
       x = "River",
       y = "Count",
       fill = "Species")
```

# Task 6

```{r}
# pie chart of species
tab_species <- with(ddt, table(SPECIES))
pie(tab_species, main = "Fish Species", col = 1:length(tab_species))

# pie chart of rivers
pie(tab_river, main = "Rivers", col = 1:length(tab_river))
```

# Task 7

```{r}
#boxplot ddt
boxplot(ddt$DDT, main = "Boxplot of DDT", ylab = "DDT", col = "lightgray")

#boxplot weight
boxplot(ddt$WEIGHT, main = "Boxplot of WEIGHT", ylab = "WEIGHT", col = "lightgray")

#boxplot length
boxplot(ddt$LENGTH, main = "Boxplot of LENGTH", ylab = "LENGTH", col = "lightgray")
```

# Task 8

```{r}
#coplot length v weight given river
coplot(LENGTH ~ WEIGHT | RIVER, data = ddt, xlab = "WEIGHT", ylab = "LENGTH")

#coplot ddt v weight given species
coplot(DDT ~ WEIGHT | SPECIES, data = ddt, xlab = "WEIGHT", ylab = "DDT")
```

# Task 9

```{r}
library(ggplot2)

#boxplot
ggplot(ddt, aes(x = SPECIES, y = WEIGHT, fill = RIVER)) +
  geom_boxplot() +
  ggtitle("SPECIES vs WEIGHT — Dastin Scott") +
  labs(x = "SPECIES", y = "WEIGHT")

#violin plot
ggplot(ddt, aes(x = RIVER, y = LENGTH, fill = SPECIES)) +
  geom_violin(trim = FALSE) +
  ggtitle("RIVER vs LENGTH — Dastin Scott") +
  labs(x = "RIVER", y = "LENGTH")

#scatter plot
ggplot(ddt, aes(x = WEIGHT, y = LENGTH, color = SPECIES)) +
  geom_point() +
  ggtitle("WEIGHT vs LENGTH — Dastin Scott") +
  labs(x = "WEIGHT", y = "LENGTH")
